Fixed-Shape Ellipse by Three Points

### 2019-10-17 ONPCE/slides.tex
wip

\usepackage{amsthm, amsfonts, amsmath, amssymb} % Pacote para a definicao dos ambientes matematicos
\usepackage[brazilian]{babel}
\usepackage[utf8]{inputenc}
\usepackage[mathscr]{eucal}
\usepackage{tikz}
\usetikzlibrary{quotes, angles, intersections}
\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\Pp}{\mathscr{P}}
\newcommand{\Cc}{\mathscr{C}}
\newcommand{\E}{\mathscr{E}}
\newcommand{\norm}[2][2]{\left\lVert#2\right\rVert_{#1}}

@@

		%\caption{The ellipse as a parametric curve.}
		\includegraphics[scale=.5]{e3psol2.pdf}
		\label{fig:e3psol2}
		\caption{Every solution for that instance.}
	\end{figure}
\end{frame}

@@

	Given any non-colinear points, there is an unique circumscribed circle. Also, we can turn an axis-parallel ellipse into a circle:
	\begin{figure}
		\centering

@@

	Let's rotate the points instead of rotating the ellipse. It is equivalent:

	\begin{figure}
		\centering
		%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
		\includegraphics{circumscribed-circle}
		\caption{}
		\label{fig:circumscribed-circle}
	\end{figure}
\end{frame}

@@

	\begin{itemize}
		\item Translate the points so $u=(0, 0)$.
		\item Scale the $x$-axis by $\frac{b}{a}$.
	\end{itemize}

	Then, we rotate the points. The rotated points are $\varphi(u, \theta), \varphi(v, \theta), \varphi(w, \theta)$:

	\begin{equation*}\label{eq:trpnts}
	\varphi(p, \theta)=\left[\begin{array}{cc}
	\frac{b}{a}&0\\
	0&1
	\end{array}\right]
	\left[\begin{array}{cc}
	\cos{\theta}&\sin{\theta}\\
	-\sin{\theta}&\cos{\theta}
	\end{array}\right]\left[\begin{array}{c}
	p_x\\
	p_y
	\end{array}\right].
	\end{equation*}

	After that, we need to look at the circumscribed circle formed by $\varphi(u, \theta), \varphi(v, \theta), \varphi(w, \theta)$.
\end{frame}

	\textbf{An angle is a solution if, and only if the circumscribed circle of the points  $\varphi(u, \theta), \varphi(v, \theta), \varphi(w, \theta)$ has radius $b$.}

	\begin{itemize}
		\item This is a problem of one variable!

	\end{itemize}

@@

	%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
	\includegraphics[scale=.5]{036}

%	\label{fig:circumscribed-circle}
\end{figure}
\end{frame}

@@

		%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
		\includegraphics[scale=.5]{012}
		\caption{Another example of $\xi$.}
		%	\label{fig:circumscribed-circle}
	\end{figure}
\end{frame}

>>>>

\usepackage{amsthm, amsfonts, amsmath, amssymb} % Pacote para a definicao dos ambientes matematicos
\usepackage[brazilian]{babel}
\usepackage[utf8]{inputenc}
\usepackage[mathscr]{eucal}
\usepackage{tikz}
\usetikzlibrary{quotes, angles, intersections}
\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\Pp}{\mathscr{P}}
\newcommand{\bigO}{\mathscr{O}}
\newcommand{\Cc}{\mathscr{C}}
\newcommand{\E}{\mathscr{E}}
\newcommand{\norm}[2][2]{\left\lVert#2\right\rVert_{#1}}

@@

		%\caption{The ellipse as a parametric curve.}
		\includegraphics[scale=.5]{e3psol2.pdf}
		\label{fig:e3psol2}
		\caption{Every solution for the instance shown previously.}
	\end{figure}
\end{frame}

@@

	An ellipse with shape $(a, b)$ can be transformed into a circle of radius $b$ through scaling the $x$-axis by $\frac{b}{a}$:
	\begin{figure}
		\centering

@@

	Let's rotate the points instead of rotating the ellipse:

	\begin{figure}
		\centering
		%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
		\includegraphics[scale=.5]{sol1}
		\caption{Three points at their initial location.}
		\label{fig:sol1}
	\end{figure}
\end{frame}

	Firstly, we rotate leaving one point fixed at $(0, 0)$:

	\begin{figure}
		\centering
		%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
		\includegraphics[scale=.5]{sol2}
		\caption{After rotation.}
		\label{fig:sol2}
	\end{figure}
\end{frame}

	Then, we scale by $\frac{b}{a}$ and check the radius of the circle:

	\begin{figure}
		\centering
		%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
		\includegraphics[scale=.5]{sol3}
		\caption{After scaling.}
		\label{fig:sol33}
	\end{figure}
\end{frame}

	If the radius is $b$, the angle is a solution:

	\begin{figure}
		\centering
		%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
		\includegraphics[scale=.5]{sol4}
		\caption{One solution for this instance.}
		\label{fig:sol4}
	\end{figure}
\end{frame}

@@

	\begin{itemize}
		\item Translate the points so $u=(0, 0)$.
		\item Rotate by $\theta$ and scale the $x$-axis by $\frac{b}{a}$.
		\item Find the $\theta$'s which produce a circle with radius $b$.
	\end{itemize}

	This transformation is expressed by:

	\begin{equation*}\label{eq:trpnts}
	\varphi(p, \theta)=\left[\begin{array}{cc}
	\frac{b}{a}&0\\
	0&1
	\end{array}\right]
	\left[\begin{array}{cc}
	\cos{\theta}&\sin{\theta}\\
	-\sin{\theta}&\cos{\theta}
	\end{array}\right]\left[\begin{array}{c}
	p_x\\
	p_y
	\end{array}\right],
	\end{equation*}
	with $p=u, v, w$.

	\textbf{This is an one variable problem on a closed interval!}

@@

	%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
	\includegraphics[scale=.5]{036}
	\caption{An example of $\xi$.}
%	\label{fig:circumscribed-circle}
\end{figure}
\end{frame}

@@

		%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
		\includegraphics[scale=.5]{012}
		\caption{Yet another example of $\xi$.}
		%	\label{fig:circumscribed-circle}
	\end{figure}
\end{frame}

\begin{frame}{Chebyshev Interpolation}

	\begin{block}{Chebyshev Polynomial}
		$T_n: [-1, 1] \mapsto [-1, 1]$ is the $n$-degree Chebyshev polynomial:
		\begin{equation*}
		T_n(\cos t) = \cos(nt)
		\end{equation*}
	\end{block}
	Also, it can be defined recursively:	
	\begin{align*}
	&T_0(x) = 1\\
	&T_1(x) = x\\
	&T_n(x) = 2xT_{n-1}(x) - T_{n-2}(x)
	\end{align*}

\begin{frame}{Chebyshev Interpolation}
	Interpolation on the roots of $T_n$, also known as Chebyshev Nodes:

	\begin{equation*}
		x_k = \cos\left(\pi\frac{2k - 1}{2n}\right)
	\end{equation*}

	The interpolation of a function $f : [-1, 1] \mapsto \R$ can be written directly using Chebyshev polynomial as basis:

	\begin{equation*}
	f(x) \approx \sum_{k=0}^{n} a_k T_k(x)
	\end{equation*}

	\begin{itemize}
		\item A simple change of coordinates lets the interpolation to be done on any closed interval!
	\end{itemize}
\end{frame}

\begin{frame}{Chebyshev Interpolation}
		Why is it good?

	\begin{itemize}
		\item Numerically stable! Way better then polynomials in the power format.

		\item No Runge's Phenomenon, the interpolation converges to $f$.
		\begin{itemize}
			\item $\bigO(n^{-m})$ if $f$ is $m$ times differentiable.
			\item $\bigO(C^n)$, for $C<1$, if $f$ is analytical in a neighborhood of $[-1, 1]$.
		\end{itemize}

		\item Very used in practice: present in external libraries like NumPy for Python.
	\end{itemize}
\end{frame}
